---
title       : Example measurement of low temperature Raman spectra
author      : Colin Bousige
date        : "`r format(Sys.time(), '%d/%m/%Y')`"
always_allow_html: yes
output:
    html_notebook: 
        toc: yes
        toc_float: TRUE
        code_folding: hide
        highlight: pygments
        number_sections: false
        theme: cosmo
bibliography: "/Users/colin/Dropbox/Biblio/all.bib" # change your path here
csl: "/Users/colin/Dropbox/Biblio/nature_notitle.csl" # change your path here
---

<!-- Edit your style here -->
<style type="text/css">
   body{font-size: 16px;font-family: Helvetica;}
   h1{font-size: 26px;color:RoyalBlue;font-weight: bold;}
   a{color:RoyalBlue;font-weight: bold;}
   h2{font-size: 22px;color:RoyalBlue;font-weight: bold;}
   h3{font-size: 18px;color:RoyalBlue;font-weight: bold;}
   code{font-size: 16px;}
</style>

<!-- Setup : load libraries, define path -->
```{r setup, include=FALSE, warning = FALSE, cache=FALSE}
# Set your own working directory :
setwd("/Users/colin/Travail/Cours/TutoR/") 
library(plotly)
library(ggplot2)
library(reshape2)
library(knitr)
library(akima)
library(grid)
library(FME)
source("myfunc.R") # this contains many functions
path <- "Data"
my_theme <- theme_bw()+
            theme(axis.text = element_text(size = 14,family = "Helvetica",colour="black"),
                  text = element_text(size = 14,family = "Helvetica"),
                  axis.ticks = element_line(colour = "black"),
                  legend.text = element_text(size = 14,family = "Helvetica",colour="black"),
                  panel.border = element_rect(colour = "black", fill=NA, size=1),
                  legend.key.height=unit(2,"cm")
                  )
```


# Goal of experiment

Write your stuff here, site some stuff [@Iijima1991; @Geim2013a].

# Measurement parameters

$~$                      | $~$
-------------------------|-----------------------------
__Spectrometer__         | LabRAM HR Evolution - Horiba
__Laser__                | 532 nm
__Grating__              | 1800 tr/mm
__Hole__                 | 100 $\mu$m
__Objective__            | $\times$ 50 LWD
__Acquisition time__     | 10 $\times$ 30 s
__Temperature control__  | Linkam TMS 92
__Position Si__          | 520.7 cm^-1^ - 1300 hits/s
__Max laser power__      | 3.2 mW
__Filter__               | 0.1 %

Blablabla


# Specters, background and fits 

Use the slider under the plot

```{r, out.width='100%', fig.asp=.85,cache=FALSE, warning=FALSE}
# Temperatures
Temp <-c(40,30,20,10,0,-10,-20,-30,-40,-50,-55,
         -60,-65,-70,-75,-80,-85,-90,-95,-100,-105,
         -110,-115,-120,-125,-130,-135,-140,-145,-150,-155)
Temp <- Temp + 273 # Celsius to Kelvin – temperature accuracy ~ ±0.3K
# Find files names
fRBM <- list.files(path=path,pattern = "RBM")
fG   <- list.files(path=path,pattern = "G")
# Create a list to store spectra for each temperature
spec <- vector("list",length=(length(fRBM)))

# Reading and storing all treated data 
for(i in 1:length(fRBM)){
# Reading raw RBM data
    dRBM <- read.table(file.path(path,fRBM[i]),sep="\t")
    xRBM <- dRBM[,1]
    dRBM <- dRBM[which(xRBM>=100 & xRBM<=300),]
    xRBM <- dRBM[,1]
    yRBM <- dRBM[,2]
    # Fitting polynomial background of order np
    backgrRBM <- baseline(xRBM,yRBM, poly=TRUE, np=0, 
                           BGpoints=list(c(100,130),c(240,300)),noise=10)$bg
    # Fitting peak positions
    if(i==1){
        params <- c(y0=0,
                    A =c(400,400,400,400),
                    x0=c(157,171,177,185),
                    FW=c(10,10,10,10))
    }
    else{params <- spec[[i-1]]$parRBM$coef}
    fitRBM <- fitspec(xRBM,yRBM-backgrRBM,params=params)
    # qplot(xRBM,yRBM-backgrRBM) +
    # geom_line(data=fitRBM$curves,aes(x,y,color=id))+theme_bw()
    
# Reading raw G data
    dG <- read.table(file.path(path,fG[i]),sep="\t")
    xG <- dG[,1]
    # dG <- dG[xG>1400,]
    xG <- dG[,1]
    yG <- dG[,2]
    # Fitting background with rubberband algorithm
    backgrG   <- baseline(xG,yG,noise=2)$bg
    # Fitting peak positions
    if(i==1) {
        params <- c(y0=0,
                    x0=c(1538,1557,1573,1594,1616),
                    A=c(900,900,900,900,900),
                    FW=c(10,10,10,10,10))
    }
    else{params <- spec[[i-1]]$parG$coef}
    fitG <- fitspec(xG[which(xG>1530)],
                    yG[which(xG>1530)]-backgrG[which(xG>1530)],
                    params=params, control=list(maxiter=1000)
                    )
    # qplot(xG,yG-backgrG) +
    # geom_line(data=fitG$curves,aes(x,y,color=id))+theme_bw()
# Storing spectrum for temperature_i and all other information
    spec[[i]] <- list(name    = paste(i," - T = ",Temp[i],sep=""),
                      T       = Temp[i],
                      visible = FALSE, 
                      RBM     = data.frame(x = xRBM,
                                           y = yRBM, 
                                           ycor = yRBM - backgrRBM, 
                                           bg = backgrRBM),
                      parRBM  = fitRBM,
                      G       = data.frame(x = xG, 
                                           y = yG, 
                                           ycor = yG - backgrG,
                                           bg = backgrG),
                      parG    = fitG
                     )
}
# For the visualization in the slider plot, spec[[1]] is the only one visible at first
spec[[1]]$visible <- TRUE
```

## RBM peak

```{r, out.width='100%', fig.asp=.85,cache=FALSE, warning=FALSE}
# Create plot with slider control
steps <- list()
p     <- plot_ly()
for (i in 1:length(fRBM)) {
   x <- spec[[i]]$RBM[,"x"]
   y <- spec[[i]]$RBM[,"y"]
   bg   <- spec[[i]]$RBM[,"bg"]
   vis  <- spec[[i]]$visible
   name <- spec[[i]]$name
   curves <- spec[[i]]$parRBM$curves
   xfit <- curves[which(curves[,"id"]=="ytot"),"x"]
   ytot <- bg + curves[which(curves[,"id"]=="ytot"),"y"]
   y1   <- bg + curves[which(curves[,"id"]=="y1"),"y"]
   y2   <- bg + curves[which(curves[,"id"]=="y2"),"y"]
   y3   <- bg + curves[which(curves[,"id"]=="y3"),"y"]
   y4   <- bg + curves[which(curves[,"id"]=="y4"),"y"]
   colors <- colorRampPalette(c("black","royalblue","seagreen","orange","red"))(4)
   p <- add_markers(p, x=x, y=y, visible = vis, 
                    name = name, type = "scatter", showlegend = FALSE, 
                    marker = list(color='black')
                    ) 
   p <- add_lines(p, x=xfit, y=ytot, visible = vis, 
                  name = name, text='Fit', showlegend = FALSE,
                  line=list(color='red', width=2)
                  )
   p <- add_lines(p, x=xfit, y=y1, visible = vis, 
                  name = name, text='Lor 1', showlegend = FALSE,
                  line=list(color=add.alpha(colors[1],.4), width=2)
                  )
   p <- add_lines(p, x=xfit, y=y2, visible = vis, 
                  name = name, text='Lor 2', showlegend = FALSE,
                  line=list(color=add.alpha(colors[2],.4), width=2)
                  )
   p <- add_lines(p, x=xfit, y=y3, visible = vis, 
                  name = name, text='Lor 3', showlegend = FALSE,
                  line=list(color=add.alpha(colors[3],.4), width=2)
                  )
   p <- add_lines(p, x=xfit, y=y4, visible = vis, 
                  name = name, text='Lor 4', showlegend = FALSE,
                  line=list(color=add.alpha(colors[4],.4), width=2)
                  )
   p <- add_lines(p, x=x, y=bg, visible = vis, 
                  name = name, text='Background', showlegend = FALSE,
                  line=list(color='royalblue', width=2)
                  )
   nlines <- 7
   step <- list(args = list('visible', rep(FALSE, length(spec)*nlines ) ), 
                label = spec[[i]]$T, 
                method = 'restyle'
                )
   step$args[[2]][1:nlines+nlines*(i-1)] <- TRUE
   steps[[i]] <- step
} 
# add slider control to plot
p <- p %>% 
  layout(sliders = list(list(x = 0, y=-.05, active = 0,
                             currentvalue = list(prefix = "Temperature: ",suffix=" K"),
                             steps = steps)),
         xaxis = list(title = "Raman Shift [1/cm]",
                     zeroline = TRUE, showline = TRUE, mirror = "ticks",linewidth = 2,
                     tickfont=list(size=14),titlefont=list(size=16)), 
         yaxis = list(title = "Intensity [arb. units]",
                     zeroline = TRUE, showline = TRUE, mirror = "ticks",linewidth = 2,
                     tickfont=list(size=14),titlefont=list(size=16))
         ) 
p
```

## G peak

```{r, out.width='100%', fig.asp=.85,cache=FALSE, warning=FALSE}
# Create plot with slider control
steps <- list()
p     <- plot_ly()
for (i in 1:length(fG)) {
   x <- spec[[i]]$G[,"x"]
   y <- spec[[i]]$G[,"y"]
   bg   <- spec[[i]]$G[,"bg"]
   vis  <- spec[[i]]$visible
   name <- spec[[i]]$name
   params <- spec[[i]]$parG$coef
   ytot <- bg + sumLor(x,params)
   y0   <- params["y0"]
   y1   <- bg + sumLor(x,c(y0,params[grep("1",names(params))]))
   y2   <- bg + sumLor(x,c(y0,params[grep("2",names(params))]))
   y3   <- bg + sumLor(x,c(y0,params[grep("3",names(params))]))
   y4   <- bg + sumLor(x,c(y0,params[grep("4",names(params))]))
   y5   <- bg + sumLor(x,c(y0,params[grep("5",names(params))]))
   colors <- colorRampPalette(c("black","royalblue","seagreen","orange","red"))(5)
   p <- add_markers(p, x=x, y=y, visible = vis, 
                    name = name, type = "scatter", showlegend = FALSE, 
                    marker = list(color='black')
                    ) 
   p <- add_lines(p, x=x, y=ytot, visible = vis, 
                  name = name, text='Fit', showlegend = FALSE,
                  line=list(color='red', width=2)
                  )
   p <- add_lines(p, x=x, y=y1, visible = vis, 
                  name = name, text='Lor 1', showlegend = FALSE,
                  line=list(color=add.alpha(colors[1],.4), width=2)
                  )
   p <- add_lines(p, x=x, y=y2, visible = vis, 
                  name = name, text='Lor 2', showlegend = FALSE,
                  line=list(color=add.alpha(colors[2],.4), width=2)
                  )
   p <- add_lines(p, x=x, y=y3, visible = vis, 
                  name = name, text='Lor 3', showlegend = FALSE,
                  line=list(color=add.alpha(colors[3],.4), width=2)
                  )
   p <- add_lines(p, x=x, y=y4, visible = vis, 
                  name = name, text='Lor 4', showlegend = FALSE,
                  line=list(color=add.alpha(colors[4],.4), width=2)
                  )
   p <- add_lines(p, x=x, y=y5, visible = vis, 
                  name = name, text='Lor 5', showlegend = FALSE,
                  line=list(color=add.alpha(colors[5],.4), width=2)
                  )
   p <- add_lines(p, x=x, y=bg, visible = vis, 
                  name = name, text='Background', showlegend = FALSE,
                  line=list(color='royalblue', width=2)
                  )
   nlines <- 8
   step <- list(args = list('visible', rep(FALSE, length(spec)*nlines ) ), 
                label = spec[[i]]$T, 
                method = 'restyle'
                )
   step$args[[2]][1:nlines+nlines*(i-1)] <- TRUE
   steps[[i]] <- step
} 
# add slider control to plot
p <- p %>% 
  layout(sliders = list(list(x = 0, y=-.05, active = 0,
                             currentvalue = list(prefix = "Temperature: ",suffix=" K"),
                             steps = steps)),
         xaxis = list(title = "Raman Shift [1/cm]",
                     zeroline = TRUE, showline = TRUE, mirror = "ticks",linewidth = 2,
                     tickfont=list(size=14),titlefont=list(size=16)), 
         yaxis = list(title = "Intensity [arb. units]",
                     zeroline = TRUE, showline = TRUE, mirror = "ticks",linewidth = 2,
                     tickfont=list(size=14),titlefont=list(size=16))
         ) 
p
```

# Temperature evolution

## RBM peak

```{r, out.width='100%', warning = FALSE, fig.asp=.85,cache=FALSE}
# Regrouping data in a 3-columns data.frame
evol <- data.frame()
for(i in 1:length(fRBM)){
    d <- data.frame(w = spec[[i]]$RBM[,"x"],
                    Intensity = spec[[i]]$RBM[,"ycor"]/max(spec[[i]]$RBM[,"ycor"]),
                    # y = spec[[i]]$RBM[,"ycor"],
                    Temperature = rep(spec[[i]]$T,length(spec[[i]]$RBM[,"ycor"]))
                    )
    evol <- rbind(evol,d)
}
# Smoothing
library(akima)
evol.interp <- with(evol, 
                    interp(x = w, y = Temperature, z = Intensity, 
                           duplicate="median",
                           xo=seq(min(evol$w), max(evol$w), length = 200),
                           yo=seq(min(evol$Temperature), max(evol$Temperature), length = 200),
                           extrap=FALSE, linear=FALSE)
                   )
# Regrouping this list to a 3-columns data.frame
melt_x <- rep(evol.interp$x, times=length(evol.interp$y))
melt_y <- rep(evol.interp$y, each=length(evol.interp$x))
melt_z <- as.vector(evol.interp$z)
evol.smooth <- na.omit(data.frame(w=melt_x, Temperature=melt_y, Intensity=melt_z))
# Plotting
breaksx <- seq(0,2000,10)
breaksy <- seq(0,400,10)
colors <- colorRampPalette(c("white","royalblue","seagreen","orange","red","brown"))(500)
PRBM <- ggplot(data=evol.smooth, aes(x=w, y=Temperature, fill=Intensity)) + 
          geom_tile() + 
          ggtitle("RBM band vs T")+
          scale_fill_gradientn(colors=colors, name='  Intensity\n[arb. units]') +
          labs(x = "Raman Shift [1/cm]",y="Temperature [K]") +
          scale_x_continuous(limits = c(130,230), expand = c(0,0), 
                             breaks = breaksx, labels = every_nth(breaksx, 2, inverse = TRUE)) + 
          scale_y_continuous(expand = c(0,0), breaks = breaksy, 
                             labels = every_nth(breaksy, 2, inverse = TRUE)) +
          my_theme + theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank()) 
ggplotly(PRBM)

parRBM   <- as.data.frame(t(recup(spec,"parRBM","coef")))
parRBM$T <- Temp
RBMc <- data.frame(T=rep(Temp,length(grep("x0",names(parRBM)))),
                   w=unlist(parRBM[grep("x0",names(parRBM))]),
                   Pic=rep(names(parRBM)[grep("x0",names(parRBM))],each=length(Temp))
                   )
colors <- colorRampPalette(c("black","royalblue","seagreen","orange","red"))(4)
PparRBM <- ggplot(data=RBMc, aes(x=T, y=w, color=Pic)) + 
           ggtitle("RBM Peak positions vs T") +
           geom_point(size=3) + 
           scale_colour_manual(values = colors) +
           labs(y = "Raman Shift [1/cm]", x="Temperature [K]") +
           my_theme 
ggplotly(PparRBM)
```

## G peak

```{r, out.width='100%', warning = FALSE, fig.asp=.85,cache=FALSE}
# Regrouping data in a 3-columns data.frame
evol <- data.frame()
for(i in 1:length(fG)){
    d <- data.frame(w = spec[[i]]$G[,"x"],
                    Intensity = spec[[i]]$G[,"ycor"]/max(spec[[i]]$G[,"ycor"]),
                    Temperature = rep(spec[[i]]$T,length(spec[[i]]$G[,"ycor"]))
                    )
    evol <- rbind(evol,d)
}
# Smoothing
library(akima)
evol.interp <- with(evol, 
                    interp(x = w, y = Temperature, z = Intensity, 
                           duplicate="median",
                           xo=seq(min(evol$w), max(evol$w), length = 200),
                           yo=seq(min(evol$Temperature), max(evol$Temperature), length = 200),
                           extrap=FALSE, linear=FALSE)
                   )
# Regrouping this list to a 3-columns data.frame
melt_x <- rep(evol.interp$x, times=length(evol.interp$y))
melt_y <- rep(evol.interp$y, each=length(evol.interp$x))
melt_z <- as.vector(evol.interp$z)
evol.smooth <- na.omit(data.frame(w=melt_x, Temperature=melt_y, Intensity=melt_z))
# Plotting
breaksx <- seq(0,2000,10)
breaksy <- seq(0,400,10)
colors <- colorRampPalette(c("white","royalblue","seagreen","orange","red","brown"))(500)
PG <- ggplot(data=evol.smooth, aes(x=w, y=Temperature, fill=Intensity)) + 
      geom_tile() + 
      # geom_hline(yintercept=283)+
      ggtitle("G band vs T")+
      scale_fill_gradientn(colors=colors, name='  Intensity\n[arb. units]') +
      labs(x = "Raman Shift [1/cm]",y="Temperature [K]") +
      scale_x_continuous(limits=c(1540,1640), expand = c(0,0), breaks = breaksx, labels = every_nth(breaksx, 2, inverse = TRUE)) + 
      # scale_x_continuous(expand = c(0,0), breaks = breaksx, labels = every_nth(breaksx, 2, inverse = TRUE)) + 
      scale_y_continuous(expand = c(0,0), breaks = breaksy, labels = every_nth(breaksy, 2, inverse = TRUE)) +
      my_theme + theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank())
ggplotly(PG)

parG   <- as.data.frame(t(recup(spec,"parG","coef")))
parG$T <- Temp
Gc <- data.frame(T=rep(Temp,length(grep("x0",names(parG)))),
                 w=unlist(parG[grep("x0",names(parG))]),
                 Pic=rep(names(parG)[grep("x0",names(parG))],each=length(Temp))
                 )
colors <- colorRampPalette(c("black","royalblue","seagreen","orange","red"))(5)
PparG  <- ggplot(data=Gc, aes(x=T, y=w, color=Pic)) + 
           ggtitle("G Peak positions vs T") +
           geom_point(size=3) + 
           scale_colour_manual(values = colors) +
           labs(y = "Raman Shift [1/cm]", x="Temperature [K]") +
           my_theme 
ggplotly(PparG)
```



```{r, out.width='100%', warning = FALSE, fig.asp=.85,cache=FALSE, include=FALSE}
# Output plots to files
PRBM <- PRBM+theme(axis.ticks.length=unit(-0.2, "cm"),
             axis.text.x = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")), 
             axis.text.y = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm"))) 
PG <- PG+theme(axis.ticks.length=unit(-0.2, "cm"),
             axis.text.x = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")), 
             axis.text.y = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm"))) 
PparG <- PparG+theme(axis.ticks.length=unit(-0.2, "cm"),
             axis.text.x = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")), 
             axis.text.y = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm"))) 
PparRBM <- PparRBM+theme(axis.ticks.length=unit(-0.2, "cm"),
             axis.text.x = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm")), 
             axis.text.y = element_text(margin=unit(c(0.5,0.5,0.5,0.5), "cm"))) 
# library(tikzDevice)
# tikz("Plots/Log.tex",height=6, width=8, pointsize = 10,standAlone=TRUE)
pdf("Plots/Log.pdf",height=6, width=8)
print(PRBM)
print(PG)
print(PparRBM)
print(PparG)
dev.off()
```

# Conclusions

Blablabla


# References


