---
date: 2017-06-07T19:56:21+02:00
title: Plotting 3D data with R
tags : ["R"]
summary : Various ways of plotting 3D datasets with R
---

In my work I do a lot of measurements that I like to plot as a color map, like the evolution of a Raman spectrum as a function of temperature, pressure or position. In some cases I end up with a 3-columns `data.frame` with x, y, and z values (_e.g._ intensity of a peak as a function of the position on my sample), in some cases I have a list of spectra evolving with a given parameter. I have been looking for the best ways to plot this kind of data as a color map, and I'd like to share it here for people who have been struggling like me to do this.

<!-- MarkdownTOC -->

- [List of spectra or 3-columns data.frame](#list-of-spectra-or-3-columns-dataframe)
- [Matrix of data](#matrix-of-data)
- [Conclusion](#conclusion)

<!-- /MarkdownTOC -->



## List of spectra or 3-columns data.frame

Let's create a dummy set of spectra that we will gather in a list. I like to do this, because it allows me to store in this list all the information I want about each spectrum, like the temperature, pressure, fit parameters, etc.

```{r, message=FALSE}
Nspec <- 40                           # Amount of spectra
T     <- seq(273,500, length=Nspec)   # Temperatures
N     <- 500                          # Size of the x vector
spec  <- vector( "list", length=Nspec) # We will store the spectra here
for (i in 1:length(spec)) {
    # Let's create dummy x and y values
    # You would normally use a read.table here 
    x <- seq(0, 100, length = N)
    # y is a double Gaussian with some noise added
    y <- runif(N,max=.1) +
         50*Gaussian(x=x, 
                     x0=c((T[i]/T[1])*20 + 40,(T[i]/T[1] -1)*25 +20 ), 
                     FWHM=c(10+runif(1,max=5),10+runif(1,max=5)))
    spec[[i]] <- list(name = paste("Spectrum",i),
                      T    = T[i],
                      data = data.frame(w=x,Intensity=y) 
                     ) # you can store other information in this list
}
# Plot two spectra to see what they look like
plot(spec[[1]]$data,pch=16,ylim=c(0,5))
points(spec[[Nspec]]$data,col="royalblue",pch=16)
legend("top",cex=1.3,bty = "n",pch=16,
    col=c("black","royalblue"),
    legend=c(spec[[1]]$name,spec[[Nspec]]$name)
    )
```
OK, so now we have some fake experimental data stored in a list called `spec`. Now we want to plot it as a color map in order to grasp the evolution of the spectra. We will do it using `ggplot`, because it's good and also because you can even make it an interactive plot thanks to `plotly`.

```{r, out.width='100%', message = FALSE, fig.asp=.7,cache=FALSE}
require(ggplot2)
require(plotly)
# Regrouping data in a 3-columns data.frame
evol <- data.frame()
for(i in 1:Nspec){
    d <- data.frame(w = spec[[i]]$data[,"w"],
                    Intensity = spec[[i]]$data[,"Intensity"],
                    T = rep(spec[[i]]$T,length(spec[[i]]$data[,"w"]))
                    )
    evol <- rbind(evol,d)
}
head(evol)
# Plotting
P1 <- ggplot(data=evol, aes(x=w, y=T, fill=Intensity)) + 
      geom_tile() +
      ggtitle("Original data") + 
      scale_fill_gradientn(colors=colorRampPalette(c("white","royalblue","seagreen","orange","red","brown"))(500),name="Intensity\n[arb. units]") +
      labs(x = "Raman Shift [1/cm]",y="Temperature [K]") +
      theme_bw()
ggplotly(P1)
```
Nice, right?  

Now let's say I want to smooth the color plot to make it nicer, or because my spectra don't have all the same sampling (because I removed some spikes or changed the spectrometer range/resolution or whatever).

```{r, out.width='100%', message = FALSE, warning = FALSE, fig.asp=.5,cache=FALSE}
# Interpolation on a regular 2D grid of size 100x100
require(akima)
evol.interp <- with(evol, 
                    interp(x = w, y = T, z = Intensity, 
                           duplicate="median",
                           xo=seq(min(evol$w), max(evol$w), length = 100),
                           yo=seq(min(evol$T), max(evol$T), length = 100),
                           extrap=FALSE, linear=FALSE)
                   )
# Regrouping this list to a 3-columns data.frame
melt_x <- rep(evol.interp$x, times=length(evol.interp$y))
melt_y <- rep(evol.interp$y, each=length(evol.interp$x))
melt_z <- as.vector(evol.interp$z)
evol.smooth <- na.omit(data.frame(w=melt_x, Temperature=melt_y, Intensity=melt_z))

# Plotting
P2 <- ggplot(data=evol.smooth, aes(x=w, y=Temperature, fill=Intensity)) + 
      geom_tile() + 
      ggtitle("Smoothed data") +
      scale_fill_gradientn(colors=colorRampPalette(c("white","royalblue","seagreen","orange","red","brown"))(500), name="Intensity\n[arb. units]") +
      labs(x = "Raman Shift [1/cm]",y="Temperature [K]") +
      theme_bw()
# ggplotly(P2)
require(cowplot)
plot_grid(P1 + theme(legend.position="none"), 
          P2 + theme(legend.position="none"),align="hv")
```


## Matrix of data

In some cases you end up with a matrix _z_, and two vectors _x_ and _y_. This is easy to plot using the base `image` function. For the sake of example, let's just use the `acast` function on our melted 3-columns data.frame:

```{r, out.width='100%', message = FALSE, warning = FALSE, fig.asp=.6,cache=FALSE}
require(reshape2)
z <- acast(evol, x~T, value.var="Intensity")
x <- unique(evol[,"w"])
y <- unique(evol[,"T"])
x <- x[order(x)]
y <- y[order(y)]
colors <- colorRampPalette(c("white","royalblue","seagreen","orange","red","brown"))(500)
par(mar=c(4, 4, .5, 4), lwd=2)
image(x,y,z,col = colors)
```
You can add a legend by using the `image.plot` function:
```{r, out.width='100%', message = FALSE, warning = FALSE, fig.asp=.6,cache=FALSE}
require(fields)
par(mar=c(4, 4, .5, 4), lwd=2)
image.plot(x,y,z,col = colors)
```
And finally, if you want to make this an interactive plot, you can use `plotly`:
```{r, out.width='100%', message = FALSE, warning = FALSE, fig.asp=.6,cache=FALSE}
require(plotly)
aX <- list(title = "Raman Shift [1/cm]")
aY <- list(title = "Temperature [K]")
# Weird but you need to use t(z) here:
z <- t(z)
# Color plot
plot_ly(x=x,y=y,z=z,type = "heatmap",colors=colors) %>% 
   layout(xaxis = aX, yaxis = aY)
```
Or, very cool, an interactive surface plot:
```{r, out.width='100%', message = FALSE, warning = FALSE, fig.asp=.9,cache=FALSE}
plot_ly(x=x,y=y,z=z,type = "surface",colors=colors) %>%
   layout(scene = list(xaxis = aX, yaxis = aY, dragmode="turntable"))
```

## Conclusion

Now you have many options to make nice color plots, just pick the one for you!